\subsection{Determinism}

{  %% chapter slide
  \setbeamercolor{background canvas}{bg=sectioncolor}
\begin{frame}{Challenge \textnumero 4}{Deterministic State Machines}

  How to assure DFAs are deterministic by construction?

  \begin{itemize}
  \item   Compute a partition of a given set of type designators,
  \item   ... even (especially) when subtype relation is unknown.
  \end{itemize}

\end{frame}
}


% Int and Int (classic case)
%% \begin{frame}{Classic Non-determinism}
%%   \centering
  
%%   \scalebox{0.95}{\input{fig-nfa}}

%%   Classic \Emph{non-determinism}: state with two exiting transitions with same label.

%% \end{frame}

\begin{frame}{Non-determinism by subtype}
  \begin{columns}[T]
    \begin{column}{0.4\textwidth}
      \centering
      
      \begin{align*}
        Int&\subseteq Number\\
        Int &\cap Number \neq \emptyset
      \end{align*}%
      \scalebox{0.8}{\input{tikz-int-double}}%
    \end{column}%
    \begin{column}{0.6\textwidth}
      \only<1>{\scalebox{0.9}{\input{fig-nfa-subtype-3}}}%
      \only<2>{\scalebox{0.8}{\input{fig-nfa-subtype-disjoint}}}%
      \only<3>{\scalebox{0.8}{\input{fig-nfa-subtype-disjoint-simple}}}%
      \only<4>{\scalebox{0.8}{\input{fig-nfa-subtype-simple}}}%
    \end{column}
  \end{columns}
  \only<3-4>{We can \Emph{decide} that state \nodecirc{3} is unreachable.}
\end{frame}



%% \newsavebox\oddbox
%% \begin{lrbox}{\oddbox}
%%   \begin{minipage}{11cm}"
%%     \input{codeOddp}
%%   \end{minipage}
%% \end{lrbox}

%% \begin{frame}{Definition of Odd type}
%%   \usebox\oddbox
%% \end{frame}


% Int and Number and Odd
% Partition into Int & Odd, Int & !Odd, !Int & Odd, !Int & !Odd
% We know Int & Odd = Odd
%         !Int & Odd = empty
%         !Int & !Odd = !Int

\begin{frame}{Non-determinism by \code{SSatisfies}}
  \begin{columns}[T]
    \begin{column}{0.35\textwidth}
      \centering
      \begin{align*}
        Odd&\subseteq Int &\text{unknown}\\
        Odd&~\cap~ !Int = \emptyset &\text{unknown}
      \end{align*}
      \scalebox{1.0}{\input{tikz-int-odd.tex}}%
    \end{column}%
    \begin{column}{0.7\textwidth}
      \only<1>{\scalebox{0.9}{\input{fig-nfa-satisfies}}}%
      \only<2>{\scalebox{0.8}{\input{fig-nfa-satisfies-intersections}}}%
      \only<3>{\scalebox{0.74}{\input{fig-nfa-satisfies-intersections-simple}}}%
    \end{column}
  \end{columns}
  \only<3>{\Emph{Unreachable} states \nodecirc{4} and \nodecirc{6}, but \Emph{undecidable}.}
\end{frame}




% Double and Odd + Double and Even


% classes A and B
% if A and B are final, then they are disjoint
% if A and B are abstract, there MIGHT be a common subclass
% if A and B are in a subtype relation, then either A &! B or B & !A is empty


\newsavebox\classbox
\begin{lrbox}{\classbox}
  \begin{minipage}{5cm}
    \input{codeClass}
  \end{minipage}
\end{lrbox}

\begin{frame}{Unsatisfiable Transitions}

  \scalebox{0.8}{\input{fig-4-unsat}}
  \begin{itemize}
  \item   If we determine a type is empty, then the transistion is \Emph{unsatifiable}.
  \item Thus we \Emph{can eliminate} the transition and unreachable states.

  \end{itemize}
\end{frame}

\begin{frame}{Indeterminant Transitions}

  \scalebox{0.8}{\input{fig-4}}

  \begin{itemize}
  \item   If we cannot determine a type is empty, the transition may
    \Emph{still be unsatisfiable}. 
  \item  However, we \Emph{cannot eliminate}    the transition and unreachable states.
  \end{itemize}
\end{frame}


\begin{frame}{Indeterminant Transitions}

  \scalebox{0.8}{\input{fig-4}}

  \begin{itemize}
    \item We \Emph{can always} determine type membership.
    \item DFAs with indeterminant transitions \Emph{correctly}
      match sequences in $O(n)$.
  \end{itemize}
\end{frame}


\begin{frame}{Non-determinism with classes}
  \begin{columns}[T]
    \begin{column}{0.45\textwidth}
      \usebox\classbox  
    \end{column}%
    \begin{column}{0.55\textwidth}
      \only<1>{\scalebox{0.95}{\input{fig-A-B}}}%
      \only<2-3>{\scalebox{0.95}{\input{fig-C-D}}}%
      \only<4>{\scalebox{0.95}{\input{fig-G-E}}}%
      \only<5>{\scalebox{0.95}{\input{fig-E-F}}}%

  \only<1>{{\LARGE \[\textcolor{blue}{A\cap B = \emptyset}\]}}%
  \only<2>{{\LARGE \[\textcolor{blue}{C\cap D = unknown}\]}}%
  \only<3>{{\LARGE \[\textcolor{blue}{C\cap D = unknown}\]}}%
  \only<4>{{\LARGE \[\textcolor{blue}{G\subset E}\]}}%
  \only<5>{{\LARGE \[\textcolor{blue}{E\cap F \neq \emptyset}\]}}%
    \end{column}%
  \end{columns}
  \only<1>{Both are \code{final}; they have no common \Emph{inhabited} subclass.}%
  \only<2>{\medskip

    We do not know whether some class inherits from both.}%
  \only<3>{\medskip

    JVM might run-time load another library which creates a common subclass of \code{C} and \code{D}}%
  \only<4>{Explicit subtype relation.}%
  \only<5>{\code{G} inherits from \code{E} and \code{F}.

    \medskip
    
    Java $>8.x$, cannot compute subclasses.

    \medskip
    
    \url{github.com/ronmamo/reflections} no longer maintained.}
\end{frame}
