\section{Determinism}

\begin{frame}{Deterministic vs non-deterministic }
  
\end{frame}

% Int and Int (classic case)
\begin{frame}{Classic Non-determinism}
  \scalebox{0.8}{\input{fig-nfa}}

  Classic \emph{non-determinism}: state with two exiting transitions with same label.

\end{frame}

\begin{frame}{Non-determinism by subtype}
  \begin{columns}
    \begin{column}{0.4\textwidth}
      \centering

    \begin{align*}
    Int&\subseteq Number\\
    Int &\cap Number \neq \emptyset
  \end{align*}
  \scalebox{0.8}{\input{tikz-int-double.tex}}
    \end{column}%
    \begin{column}{0.6\textwidth}
      \only<1>{\scalebox{0.65}{\input{fig-nfa-subtype}}}%
      \only<2>{\scalebox{0.65}{\input{fig-nfa-subtype-disjoint}}}%
      \only<3>{\scalebox{0.65}{\input{fig-nfa-subtype-disjoint-simple}}}%
      \only<4>{\scalebox{0.65}{\input{fig-nfa-subtype-simple}}}%
    \end{column}

  \end{columns}
  \only<3-4>{We can \emph{decide} that state \nodecirc{3} is unreachable.}
\end{frame}



\newsavebox\oddbox
\begin{lrbox}{\oddbox}
  \begin{minipage}{11cm}
    \input{codeOddp}
  \end{minipage}
\end{lrbox}

\begin{frame}{Definition of Odd type}
  \usebox\oddbox
\end{frame}


% Int and Number and Odd
% Partition into Int & Odd, Int & !Odd, !Int & Odd, !Int & !Odd
% We know Int & Odd = Odd
%         !Int & Odd = empty
%         !Int & !Odd = !Int

\begin{frame}{Non-determinism by \code{SSatisfies}}
  \begin{columns}
    \begin{column}{0.35\textwidth}
      \centering
      
      \begin{align*}
        Int&\subseteq Number\\
        Odd&\subseteq Int &\text{unknown}\\
        Odd&~\cap~ !Int = \emptyset &\text{unknown}
      \end{align*}
      \scalebox{0.8}{\input{tikz-int-odd.tex}}
    \end{column}%
    \begin{column}{0.7\textwidth}
      \only<1>{\scalebox{0.6}{\input{fig-nfa-satisfies}}}%
      \only<2>{\scalebox{0.6}{\input{fig-nfa-satisfies-intersections}}}%
      \only<3>{\scalebox{0.6}{\input{fig-nfa-satisfies-intersections-simple}}}%
    \end{column}
  \end{columns}
  \only<3>{States \nodecirc{4} and \nodecirc{6} are unreachable, but we cannot \emph{decide} this \emph{unsatisfiability}.}

\end{frame}




% Double and Odd + Double and Even


% classes A and B
% if A and B are final, then they are disjoint
% if A and B are abstract, there MIGHT be a common subclass
% if A and B are in a subtype relation, then either A &! B or B & !A is empty


\newsavebox\classbox
\begin{lrbox}{\classbox}
  \begin{minipage}{5cm}
    \input{codeClass}
  \end{minipage}
\end{lrbox}

\begin{frame}{Indeterminant Transitions}

  \scalebox{0.8}{\input{fig-4}}

  If we determine a type is empty, we eliminate the
  transition and unreachable states.
\end{frame}


\begin{frame}{Indeterminant Transitions}

  \scalebox{0.8}{\input{fig-4}}

  At run-time:
  \begin{itemize}

    \item We can always determine type membership.

    \item DFAs even with indeterminant transitions \emph{correctly} match sequences in linear time.  
  \end{itemize}
\end{frame}


\begin{frame}{Non-determinism with classes}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \usebox\classbox  
    \end{column}%
    \begin{column}{0.5\textwidth}
      \only<1>{\scalebox{0.8}{\input{fig-A-B}}}%
      \only<2-3>{\scalebox{0.8}{\input{fig-C-D}}}%
      \only<4>{\scalebox{0.8}{\input{fig-G-E}}}%
      \only<5>{\scalebox{0.8}{\input{fig-E-F}}}%

  \only<1>{{\LARGE \[\textcolor{blue}{A\cap B = \emptyset}\]} Both are \code{final}; they have no common \emph{inhabited} subclass.}%
  \only<2>{{\LARGE \[\textcolor{blue}{C\cap D = unknown}\]} We do not know whether some class inherits from both.}%
  \only<3>{{\LARGE \[\textcolor{blue}{C\cap D = unknown}\]} JVM might run-time load another library which creates a common subclass of \code{C} and \code{D}}
  \only<4>{{\LARGE \[\textcolor{blue}{G\subset E}\]} Explicit subtype relation.}
  \only<5>{{\LARGE \[\textcolor{blue}{E\cap F \neq \emptyset}\]} \code{G} inherits from \code{E} and \code{F}.

    \medskip
    
    Java $>8.x$, cannot compute subclasses.

    \medskip
    
    \url{github.com/ronmamo/reflections} no longer maintained.}

    \end{column}%
  \end{columns}
\end{frame}
